---
title: "GSS challange More Data"
author: "Israel Valdovinos"
date: "Fall 2025"
output: html_document
---







```{r include=FALSE,echo=FALSE}
source(file="fisherXY.R")
require(tidyverse)
require(tigerstats)
require(knitr)
require(foreign)
GSS <-read.dta(file="GSS2016.dta")
GSS <- droplevels(GSS)
mydata <- select(GSS,race,cappun)
mydata <- na.omit(mydata)
```

# Introduction

<!--In this section you explain what you are trying to show.  Where did the data come from?  What is the research or other question you are trying to answer?.  If you skimp in this section expect to be marked down in your grade. Paint a picture of the project a couple of paragraphs is expected!-->

Is attitude toward capitol punishment (cappun) independent of race (race)?
If not, then whose view differs from whom??

# Methods

<!--Decide on your methods:  use "variable analysis" or other appropriate descriptors.  Make sure to choose at least one graphical method and at least one numerical method. Discuss what data manipulations may be needed for example subsetting or other binning procedures and what sort of graphical displays and inferential tools will be used!-->

This problem was a categorical categorical problem. We found out in the categorical of race, there was a factor with 3 levels. These were called "white", "black", and "others". cappun is a categorical variable with 2 levels which are in favor of capital punishment or oppose capital punishment. These were called "favor" and "oppose"

# Results

<!--Divide this section into two sub-sections:  One for your descriptive  results and one for your inferential results.!-->

## Descriptive Results

### Graphical Descriptive Results

<!--Graphical results here.  Make sure to show your code.  Provide appropriate labels for axes, giving units if possible, and provide a good title for the graph, too.  Use the graphical results to describe the patterns if any that exist in the data as focused toward the research question. Bottom line is to explain the features of the graph that are relevant to conforming or departing to what is expected by your null hypothesis !-->

```{r}
dd3 <- mydata %>% group_by(race,cappun) %>% summarize(count=n()) %>% mutate(prcnt=count/sum(count))


# the group_by followed by summarize(count=n())

basicC <- ggplot(dd3,aes(x=race,y=count,fill=cappun))
basicC + geom_bar(stat="identity",position="dodge")

#Now for percentage plot

basicCC <- ggplot(dd3,aes(x=race,y=prcnt*100,fill=cappun)) 
basicCC + geom_bar(stat="identity", position = "dodge")
```

The white branch and the other branch are similar when it comes to capital punishment, because they support it. The black branch does not support capital punishment. When you put the blue branch on each red branch for black, white, and other, it adds up to 100.

## Now with The Most data (fewer variables gives more data points)



### Numerical Descriptive Results

<!--Numerical results go here. Use the numerical results to describe the patterns if any that exist in the data as focused toward the research question!-->

```{r}
table2 <- xtabs(~cappun + race, data=mydata)
rowPerc(table2)
colPerc(table2)
```

Of the people who favor capitol punishment almost 80% of the, are white people. Black People made up 11.72% and other made up 8.47respectively. Of those who opposed capitol punishment , 65.38%,24.11%10.51% were 'white","black"and "other" respectively.

## Now with More data



## Inferential Results

<!--State hypothesis clearly.  Make sure your discussion of the inferential test covers all the aspects that the test output produces, such as test statistic, p-value etc.  Make a decision about the null hypothesis, explain the assumptions on which the selected test/procedure was based, and why the chosen procedure satisfys the assumptions and is appropriate to answer the research question!-->

`


## More data version

```{r}
fisherXY(mydata$race,mydata$cappun)
```
Statistically, every group is different than every group. With a low p-value for each of the groups.


# Discussion and Conclusion

<!--Discussion and conclusion here.  If you found a relationship be sure to consider whether the relationship occurs because one of the variavbles causes the other, or whether they perhasps are related for some other reason.  Watch the chapter 6 videos from the GeorgeTown videos collection.!-->

Compared to the number2 branch, the master branch included many more observations—2,867 in total, about three times as many. This is because we found a statistical difference across all groups, which led to more variation in the p-values. Having more data provided stronger evidence, which was confirmed by Fisher’s test in the master branch, showing a very low p-value of 0.000017. This indicates a significant difference between Black, White, and Other respondents. The graphical results showed that White and Other groups were similar in their support for capital punishment, while the Black group was less supportive. The numerical results supported this pattern: of those who favored capital punishment, nearly 80% were White, 11.72% were Black, and 8.47% were Other. On the opposing side, 65.38% were White, 24.11% were Black, and 10.51% were Other. Overall, the results suggest that views on capital punishment vary by race, with White respondents showing the strongest support.
